---
title: "FFT Sandbox"
author: "Ryszard Czerminski"
date: "September 2, 2017"
header-includes:
   - \usepackage{amsmath}
   - \usepackage{mathtools}
   - \DeclareMathOperator*{\E}{\mathbb{E}}
   - \usepackage{hyperref}
   - \hypersetup{
      colorlinks=true,
      linkcolor=blue,
      filecolor=magenta,
      urlcolor=cyan,
    }

output:
    pdf_document
---

\section{Intro}

The goal of this document is to provide (very) brief introduction to FFT in R

\section{FFT basics}

Fourier Transform for one dimensional function of time ($f(t)$) is defined as:
$$
    \hat f(\nu) = \int_{-\infty}^{+\infty} f(t) e^{-2\pi i \nu t} dt
$$
Note that for real valued function \href{https://en.wikipedia.org/wiki/Fast_Fourier_transform#FFT_algorithms_specialized_for_real_and.2For_symmetric_data}{the power spectrum is symmetric} 

\section {Function definitions}
```{r echo=TRUE}
gen.signal <- function(
                   amplitude, # amplitudes
                   period,    # periods
                   phase,     # phases
                   time,      # sampling times
                   shift=0,   # constant added to the periodic sinusoidal signal
                   noise=0,   # how much noise to add
                   verbose=0) {
  n <- unique(c(length(amplitude), length(period), length(phase)))
  if (length(n) != 1) stop('amplitude, periods, and phase lengths must be equal')
  y <- rep(shift, length(t))
  for (i in 1:n) y <- y + amplitude[i] * sin(phase[i] + 2*pi*time/period[i])
  if (noise > 0) y <- y + noise * rnorm(length(y), sd=sd(y))
  return (y)
}
get.fft <- function(y,             # signal
                    t,             # sampling times
                    mean.rm=FALSE) # keep/remove mean of the signal
{
  dt <- t[2] - t[1]
  df <- 1/(max(t) - min(t))
  fnyq <- 1/(2*dt)           # Nyquist frequency
  f <- seq(0, 2*fnyq, by=df) # frequencies for FFT output (in Hz)
  if (mean.rm) y <- y - mean(y)
  ft <- 2 * fft(y)/length(y) # normalized fourier transform
  pw <- abs(ft) # power spectrum
  return (list(f=f, ft=ft, pw=pw))
}
plt.power <- function(y,        # signal
                      t,        # sampling times
                      pw,       # power spectrum
                      f,        # frequencies
                      show='',  # what to display
                      eps=1e-6, # small number to ignore floating-point errors
                      ext=1.1)  # how much to extend the plot beyond the last maximum
  {
    n <- unique(c(length(y), length(t), length(f), length(pw)))
    if (length(n) != 1) stop('y,t,pw,f lengths should be equal')
    if (show == 'all') {
      mx <- n
    } else if (show == 'half') {
      mx <- n/2
    } else {
      z <- diff(pw[1:(n/2)], differences=2)
      w <- which(z < -eps)
      mx <- min(max(w) * ext, n/2) # index for the highest frequency maximum
    }
    idx <- 1:mx
    par(mfrow=c(1,2))
    plot(t, y, type='l', xlab='time [s]', ylab='signal value', main='original signal')
    plot(f[idx], pw[idx], type='l', xlab='freq [Hz]', ylab='power', main='FFT output')
    par(mfrow=c(1,1))
}
```
```{r fig.height=4, fig.width=8}
amplitudes <- c(1, 1/2, 1/4, 1/8)
periods <- c(1, 1/2, 1/4, 1/8)
phases <- runif(4, 0, 2*pi)
dt <- 1/64
nsamples <- 4 * (1/dt)
t <- (0:(nsamples-1)) * dt
y <- gen.signal(amplitudes, periods, phases, t, shift=0, verbose=1)
v <- get.fft(y, t)
plt.power(y, t, v$pw, v$f)
```


```{r fig.height=4, fig.width=8}
y <- gen.signal(amplitudes, periods, phases, t, shift=1, noise=0.1, verbose=1)
v <- get.fft(y, t, mean.rm=FALSE)
plt.power(y, t, v$pw, v$f)
```
and that power component at $\nu = 0$ can be removed simply by subtracting the mean from the signal.
```{r fig.height=4, fig.width=8}
y <- gen.signal(amplitudes, periods, phases, t, shift=1, verbose=1)
v <- get.fft(y, t, mean.rm=TRUE)
plt.power(y, t, v$pw, v$f)
```
\section{Some points to note}
\begin{enumerate}
\item phases do not have influence on the power spectrum - note that when you repeat \emph{knit}
      process phases are generated randomly and the signal looks different each time,
      but the power spectrum does not change.
      Phases influence however (complex) Fourier Transform values.
\item power spectrum value at $\nu = 0$ reflects mean value of the signal.
\item sharpness (resolution) of the spectrum is determined by the length of the signal -
      experiment with \texttt{nsamples} to see this.
\item you may also want to play with \texttt{noise} parameter to see how adding random signal
      corrupts the power spectrum
\item FFT is fastest if the number of points in the signal to analyze is equal to power of two,
      e.g. $2^9 = 512$, $2^{10} = 1024$, etc.; see help(fft)

\end{enumerate}

\section{References}
\begin{enumerate}
\item \href{https://en.wikipedia.org/wiki/Fourier_transform}{Fourier transform}
\item \href{https://en.wikipedia.org/wiki/Discrete_Fourier_transform}{Discrete Fourier transform}
\item \href{https://en.wikipedia.org/wiki/Fast_Fourier_transform}{FFT}
\item \href{https://en.wikipedia.org/wiki/Nyquist_frequency}{Nyquist frequency}
\end{enumerate}
